AppPAL for Android

### 2015-09-20 paper.tex
updates

\begin{figure}
  \newcommand{\bracetext}[1]{\text{\sffamily #1}}
  \newcommand{\smalltext}[1]{\text{\ttfamily\footnotesize #1}}
  \centering
  \begin{equation*}\small
    \begin{array}{r l}\footnotesize
      \overbrace{\smalltext{"user"}}^{\bracetext{speaker}} &
      \smalltext{ says }\overbrace{\overbrace{\smalltext{ App }}^{\bracetext{subject}}\overbrace{\smalltext{ isRunnable}}^{\bracetext{predicate}}}^{\bracetext{fact}} \\
      & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
      & \overbrace{\smalltext{ where hasPermission(App, "INTERNET") = true}}^{\bracetext{constraint}}.
    \end{array}
  \end{equation*}
  \caption{Structure of an AppPAL assertion.}
\label{fig:assertion}
\end{figure}

@@

Like Becker~\etal~we make use of an assertion context to store known statements.
We also cache intermediate results, and ground facts to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.
Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
  it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.

@@

\begin{minipage}[b]{0.49\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
def cond(ac, rt, q, d)
  ac.add q.fetch if q.isFetchable
  ac.assertions.each do |a|
    if (u = q.unify a.consequent) &&
       (a = u.substitution a).variables == none
      return checkConditions ac, rt, a, d
  return Failure
\end{lstlisting}
\end{minipage}
\begin{minipage}[b]{0.49\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
def checkConditions(ac, rt, a, d)
  getVarSubs(a,ac.constants).each do |s|
    sa = s.sub a
    if sa.antecedents.all
        { |a| evaluate(ac, rt, a, d).isValid }
      p = evaluateC sa.constraint
      return Proven if p.isValid
  return Failure
\end{lstlisting}
\end{minipage}
\caption{Partial-pseudocode for AppPAL evaluation.}
\label{fig:pseudocode}
\end{figure}

@@

We envisage AppPAL running on a mobile phone checking apps.
For it to be a viable solution the policy checking performance must be reasonable: capable of checking apps against a policy as quickly as possible.
The policy checking search procedure is at its slowest when performing large repeated delegations.
Two synthetic benchmarks were created to check that the search procedure performed acceptably.
Each benchmark consisted of a repeated chain of delegations.
The \emph{straight} benchmark consists of a single long chain of delegation.
The \emph{forking} benchmark consists of a binary tree of principals delegating to each other.
These benchmarks are reasonable as they model the worst kinds of policies to evaluate---though worse ones could be designed by forking even more.
A snippet of each policy is shown in \autoref{fig:policy-snippet}.

@@

\begin{figure}\centering
  \begin{minipage}{0.49\linewidth}
    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
    '1' says '2' can-say App isInstallable.
    '2' says '3' can-say App isInstallable.
    '3' says '4' can-say App isInstallable.
    '4' says '5' can-say App isInstallable.
    '5' says '6' can-say App isInstallable.
    '6' says '7' can-say App isInstallable.
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
    '1' says '2' can-say App isInstallable.
    '1' says '3' can-say App isInstallable.
    '2' says '4' can-say App isInstallable.
    '2' says '5' can-say App isInstallable.
    '3' says '6' can-say App isInstallable.
    '3' says '7' can-say App isInstallable.
    \end{lstlisting}
  \end{minipage}
  \caption{Straight and forking policies used for benchmarking.}
  \label{fig:policy-snippet}
\end{figure}

\begin{figure}
  \begin{minipage}{0.49\linewidth}
    \begin{tabular}{c c c c}
       \toprule
       Policy   & Nodes & Parse (s) & Check (s) \\
       \midrule
       Straight & 10    & 0.01      & 0.04      \\
       Straight & 100   & 0.05      & 1.00      \\
       Straight & 500   & 0.32      & 21.07     \\
       Straight & 1000  & 0.43      & 87.75     \\
       \midrule
       Forking  & 256   & 0.10      & 1.93      \\
       Forking  & 512   & 0.23      & 7.47      \\
       Forking  & 1024  & 0.45      & 28.29     \\
       Forking  & 2048  & 0.85      & 117.68    \\
       \bottomrule
    \end{tabular}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
    \includegraphics[width=\linewidth]{figures/benchmarks.eps}
  \end{minipage}
  \caption{Benchmarking results on a Nexus 4 Android phone.}
  \label{fig:benchmarks}
\end{figure}

@@

Throughout we have asserted that user's have policies and that there is a need for policy enforcement tools.
Corporate mobile security \ac{BYOD policies} have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq}.
Using app installation data from Carat~\cite{Oliner:2013ht} we used AppPAL to find the apps satisfying each policy Lin~\etal~identified and measured the extent each user was following them.

@@

Using AppPAL we have written short policies describing user behaviour and used these to identify the users following them to varying degrees.
Some limitations of our results include:
\begin{itemize}
\item
  We do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
  So a user may have apps installed that break the policy without us knowing.
\item
  Recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
  Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may be more conservative than our analysis suggests.
\item
  The AppPAL policies we used are a simplification of the identified privacy preferences.
  User's could be following policies more nuanced than our implementation.
  We believe our approximation is valid, however, as users will not know the precise reason an app uses a permission, and may decline to install it anyway.
  Our policies could be made more precise by incorporating the app category, and allowing apps to have a permission when it is appropriate (i.e.~a text messaging client will need the \texttt{SEND_SMS} permission to run).
\end{itemize}

>>>>

\begin{figure}
  \newcommand{\bracetext}[1]{\text{\sffamily #1}}
  \newcommand{\smalltext}[1]{\text{\ttfamily\scriptsize #1}}
  \centering
  \begin{minipage}{0.49\linewidth}
    \begin{equation*}\small
      \begin{array}{r l}\footnotesize
        \overbrace{\smalltext{`user'}}^{\bracetext{speaker}} &
        \smalltext{ says }\overbrace{\overbrace{\smalltext{ App }}^{\bracetext{subject}}\overbrace{\smalltext{ isRunnable}}^{\bracetext{predicate}}}^{\bracetext{fact}} \\
        & \overbrace{\smalltext{ if App isFree}}^{\bracetext{condition}} \\
        & \overbrace{\smalltext{ where hasPermission(App, `INTERNET') = true}}^{\bracetext{constraint}}.
      \end{array}
    \end{equation*}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
    \begin{lstlisting}[basicstyle=\ttfamily\scriptsize,keywordstyle=\ttfamily\scriptsize]
      Assertion := <E> says <Fact>
                     (if (<Fact>,)+)?
                     (where <Constraint>)?.
      Fact := <E> (isRunable | ...)
            | <E> can-say (0 | inf) <Fact>
            | <E> can-act-as <E>
      E := Variable | `constant'
    \end{lstlisting}
  \end{minipage}
  \caption{Structure and simplified grammar of an AppPAL assertion.}
  \label{fig:assertion}
\end{figure}

@@

Like Becker~\etal~we make use of an assertion context to store known statements.
We also cache intermediate results, and ground facts to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.
% Similarly when we use a statement from \emph{Emma} that \emph{Google-Play can-say} whether an app is buyable;
%   it is sensible to go fetch from the store whether the app is saleable and make Google say it then and there.

@@

\begin{minipage}[b]{0.49\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
def cond(ac, rt, q, d)
  ac.add q.fetch if q.isFetchable
  ac.assertions.each do |a|
    if (u = q.unify a.consequent) &&
       (a = u.sub a).variables == none
      return checkConditions ac, rt, a, d
  return Failure
\end{lstlisting}
\end{minipage}
\begin{minipage}[b]{0.49\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
def checkConditions(ac, rt, a, d)
  getVarSubs(a,ac.constants).each do |s|
    sa = s.sub a
    if sa.antecedents.all
        { |a| evaluate(ac, rt, a, d).isValid }
      p = evaluateC sa.constraint
      return Proven if p.isValid
  return Failure
\end{lstlisting}
\end{minipage}
\caption{Partial-pseudocode for AppPAL evaluation.}
\label{fig:pseudocode}
\end{figure}

@@

We envisage AppPAL running on a mobile phone checking apps when they are installed.
Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
Downloading and installing an app takes roughly 30s on an Android phone.
If checking a policy delays this even further user's may become annoyed and disable AppPAL.

The policy checking search procedure is at its slowest when having to delegate repeatedly;
  with the depth of the delegation tree being the biggest factor for slowing the search.
Synthetic benchmarks were created to check that the search procedure performed acceptably.
Each benchmark consisted of a chain of delegations.
The \emph{1 to 1} benchmark consists of a single long chain of delegation.
The \emph{1 to 2} benchmark consists of a binary tree of principals delegating to each other; where each principal delegated to 2 others.
Similarly the \emph{1 to 3} benchmark consists of a tree where each principal delegated to 3 others.
These benchmarks are reasonable as they model the worst kinds of policies to evaluate---though worse ones could be designed by delegating even more.
% A snippet of the \emph{1 to 1} and \emph{1 to 3} policies is shown in \autoref{fig:policy-snippet}.
For each benchmark we controlled the number of principals in the policy file:
  as the number of principals increased so did the size of the policy.

@@

% \begin{figure}\centering
%   \begin{minipage}{0.49\linewidth}
%     \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
%     '1' says '2' can-say App isInstallable.
%     '2' says '3' can-say App isInstallable.
%     '3' says '4' can-say App isInstallable.
%     '4' says '5' can-say App isInstallable.
%     '5' says '6' can-say App isInstallable.
%     '6' says '7' can-say App isInstallable.
%     \end{lstlisting}
%   \end{minipage}
%   \begin{minipage}{0.49\linewidth}
%     \begin{lstlisting}[basicstyle=\ttfamily\scriptsize, keywordstyle=\ttfamily\slshape\scriptsize]
%     '1' says '2' can-say App isInstallable.
%     '1' says '3' can-say App isInstallable.
%     '2' says '4' can-say App isInstallable.
%     '2' says '5' can-say App isInstallable.
%     '3' says '6' can-say App isInstallable.
%     '3' says '7' can-say App isInstallable.
%     \end{lstlisting}
%   \end{minipage}
%   \caption{Snippets of the \emph{1 to 1} and \emph{1 to 3} policies used for benchmarking.}
%   \label{fig:policy-snippet}
% \end{figure}

\begin{figure}
  \begin{minipage}{0.40\linewidth}
    \scriptsize
    \begin{tabular*}{0.80\linewidth}{c c r@{.}l}
      \toprule
      Delegations & Principals & \multicolumn{2}{c}{Time (s)} \\
      \midrule
      1 to 1 & 10   &  0&01 \\
      1 to 1 & 100  &  1&00 \\
      1 to 1 & 500  & 20&90 \\
      1 to 1 & 1000 & 88&73 \\
      \midrule
      1 to 2 & 10   &  0&01 \\
      1 to 2 & 100  &  0&43 \\
      1 to 2 & 500  &  7&36 \\
      1 to 2 & 1000 & 27&47 \\
      \midrule
      1 to 3 & 10   &  0&01 \\
      1 to 3 & 100  &  0&24 \\
      1 to 3 & 500  &  3&99 \\
      1 to 3 & 1000 & 15&28 \\
      \bottomrule
    \end{tabular*}
  \end{minipage}
  \begin{minipage}{0.60\linewidth}
    \includegraphics[width=\linewidth]{figures/benchmarks.eps}
  \end{minipage}
  \caption{Benchmarking results on a Nexus 4 Android phone.}
  \label{fig:benchmarks}
\end{figure}

@@

Throughout we have asserted that user's have policies and that there is a need for policy enforcement tools.
Corporate mobile security \ac{byod} policies have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq}.
Using app installation data from Carat~\cite{Oliner:2013ht} we used AppPAL to find the apps satisfying each policy Lin~\etal~identified and measured the extent each user was following them.

@@

Using AppPAL we have written short policies describing user behaviour and used these to identify the users following them to varying degrees.
Some limitations of our results include:
\begin{itemize}
\item
  We do not have the full user purchase history, and we can only find out about apps whose names match those in available databases.
  So a user may have apps installed that break the policy without us knowing.
\item
  Recently downloaded apps used for experiment may not be the same version that users had, in particular, their permissions may differ.
  Permissions tend to increase in apps over time~\cite{Wei:2012id}; so a user may be more conservative than our analysis suggests.
\item
  The AppPAL policies we used are a simplification of the identified privacy preferences.
  User's could be following policies more nuanced than our implementation.
  We believe our approximation is valid, however, as users will not know the precise reason an app uses a permission, and may decline to install it anyway.
  Our policies could be made more precise by incorporating the app category, and allowing apps to have a permission when it is appropriate (i.e.~a text messaging client will need the \texttt{SEND\_SMS} permission to run).
\end{itemize}
